Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-07-09 draft.tex
Major update, as discussed with Fred

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
%% Shortcuts %%
\newcommand{\dd}{\text{d}}
\newcommand{\refapp}[1]{app.~(\ref{app:#1})}
\newcommand{\refeq}[1]{eq.~(\ref{eq:#1})}
\newcommand{\refsec}[1]{sec.~(\ref{sec:#1})}
\newcommand{\reftab}[1]{tab.~(\ref{tab:#1})}
\newcommand{\nuvec}{\vec{\nu}}
\newcommand{\thvec}{\vec{\vartheta}}
\let\oldtheta\theta
\renewcommand{\theta}{\vartheta}
%\newcommand{\GeV}{\text{GeV}}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}
\let\eps\varepsilon
\newcommand{\aver}[1]{\langle #1 \rangle}
\newcommand{\what}[1]{\widehat{#1}}
\newcommand{\wwhat}[1]{\widehat{\widehat{#1}}}
\DeclareMathOperator{\cov}{Cov}

@@

@@

\preprint{SI-HEP-2014-xx}
\title{Sample-Based Determination of Angular Observables}
\author{Frederik Beaujean}
\email{frederik.beaujean@lmu.de}
\affiliation{C2PAP, Universe Cluster, Ludwig-Maximilians-Universit\"at M\"unchen, Garching, Germany}
\author{Marcin Chrzaszcz}
\email{mchrzasz@cern.ch}
\author{Nicola Serra}
\email{nicola.serra@cern.ch}
\affiliation{Physik-Institut, Universit\"at Z\"urich, Z\"urich, Switzerland}
\author{Danny van Dyk}
\email{vandyk@tp1.physik.uni-siegen.de}
\affiliation{Theoretische Physik 1, Naturwissenschaftlich-Technische Fakult\"at,
Universit\"at Siegen, Siegen, Germany}

\begin{abstract}
In this letter we study how to obtain a set of angular observables
$\{P_i\}$ that arise in a generic multi-body process without the need
to carry out a likelihood fit of an angular distribution to the
measured events. Instead, we present a method that only relies on
orthogonality of angular functions, and estimation of integrals
by means of Monte Carlo techniques.
\end{abstract}

@@

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the rare decay FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
he method we describe in the remainder of this letter is general and applies to arbitrary
decay or scattering processes.\\

Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $P_i$ as a coefficient in the probability density
function (PDF) of the process,
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) = \sum_i P_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$f_i(\thvec)$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega f_i(\thvec) f_j(\thvec) \dd^N \theta = \delta_{ij}\,.
\end{equation}
For the purpose of this letter, however, it suffices that the system
of angular functions $f_i(\thvec)$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays, the transformations
are listed in the appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.  Note that $P_i$ is \emph{not} a PDF.  \\

For particle decays, $P$ is generally expressed in terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\dd^{N+M}\Gamma}{\dd \nu_1 \dots \dd \nu_M\, \dd \theta_1 \dots \theta_N}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\dd^{N+M}\sigma}{\dd \nu_1 \dots \dd \nu_M\, \dd \theta_1 \dots \theta_N}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process.\fred{How would $P$ remain a proper probability such that
  $\int P = 1$? You can't simply change the normalization. I propose a
  change of notation: I find it terribly confusing that $P(\nuvec,
  \thvec)$ is a probability but $P_i(\nuvec)$ is not. For $P_i$, later
  on the symbols $J_i, K_i$ etc are used. We don't have to use the
  same symbol everywhere but let's not use $P_i$}\\

In the remainder of this letter we discuss how to obtain the angular
observables $P_i(\nuvec)$ in an experimental setup where each recorded
event is (approximately) distributed according to $P$.  We establish
the statistical basics in section \ref{sec:sample-based-det}. Section
\ref{sec:systematics} is dedicated to the impact of systematic effects
such as mismodeling the underlying physics or detector acceptance
effects. Numerical studies for one uni-angular and one triple-angular
distribution are provided in section \ref{sec:numerics}. In a series
of appendices we provide the orthonormal bases of angular functions
relevant to several rare $b$ decays.

@@

The orthonormality relations \refeq{def-ortho-rel} imply that a single angular observable $P_i$
can be projected out of the full PDF $P$ by means of
\begin{equation}
    \label{eq:det-Pi-analytical}
    P_i(\nuvec) = \int_{\thvec = 0}^{2\pi} f_i(\thvec) P(\nuvec, \thvec) \dd^N \theta\,.
\end{equation}
In point of fact, $P_i$ is the $f_i$-th moment of the PDF $P$\fred{This
  notation is uncommon. Is that how we \emph{define} the $f_i$-th
  moment?}. Additionally, integration over the non-angular variables
yields
\begin{equation}
    \langle P_i\rangle
    \equiv \int P_i(\nuvec) \dd^M \nu
    = \int \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\,.
\end{equation}
The remainder of this section  describes the method of moments, in which we replace the
analytical integration \refeq{det-Pi-analytical} by Monte Carlo (MC) estimators.\\

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int g(x) P(x) \dd x
\end{equation}
can be approximated \cite{MCSM-2004} by an MC estimator $\what{E_P[g]}$
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1} g(x^{(k)})
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$; i.e.,
\begin{equation}
    x^{(k)} \sim P\,.
\end{equation}
Throughout this letter we denote all MC estimators with a wide hat.\\

We assume, for a moment, that $P$ describes the distribution of the underlying
physical process truthfully. In that case, all events detected in an experimental setup are in fact, up to
detection efficiencies, distributed as $P$. \fred{Paragraph is redundant an can be omitted.}

Application of \refeq{mc-id} then yields \fred{notation is very confusing $P_i$ for $x^{(k)} \sim P$}
\begin{equation}
    \langle P_i\rangle \to \widehat{\langle P_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} f_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
P_i\rangle$
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle P_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} P_i(\nuvec) \dd^M \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\\
    & = \int \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        %% \theta^{(m)}(\nuvec - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec)
        \dd^N\theta\right]\dd^M \nu\,,
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{bin-importance}
    \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} f_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This basically reduces the estimation of
$\aver{P_i}_{\vec{a},\vec{b}}$ to a (weighted) counting experiment. In
the limit $N \to \infty$, the random vector $( \dots,
\aver{P_i}_{\vec{a},\vec{b}}, \dots)$ follows a multivariate Gaussian
distribution \citeneeded\fred{Why?  multivariate in $\widehat{P_i}$?}
with a mean vector estimated by \refeq{bin-importance} and covariance
estimated as
\begin{equation}
    \cov[\aver{P_i},\aver{P_j}]_{\vec{a},\vec{b}} \to \what{\cov}[\aver{P}_i, \aver{P}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[{f}_i(x^{(k)}) - \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[f_j(x^{(k)}) - \widehat{\langle P_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly \fred{So what value of $K$ is close enought to $\infty$?} converge towards the true covariance matrix.
This will allow to construct constraints on optimized observables with correlated uncertainties
from the experimental constraints on the angular observables. (See \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf}
for definitions of such optimized observables in $B\to K^*\ell^+\ell^-$ decays).

@@

In \refsec{sample-based-det}, we assume that the PDF $P$ describe the physical reality accurately,
and that the experiment observes each event with perfect accuracy. In order to estimate systematic
uncertainties, we lift these assumptions.

@@

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
We will now show that this is the case for the interesting class of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$,
which includes the rare $b\to s$-mediated $B$ decay $B \to K\pi\ell^+\ell^-$ and the $V_{ub}$ suppressed decay
$B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in the small-width approximation
and when assuming a pure $P$-wave, resonant final state. An extension to $S-P$ interference has been
studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012mb,Becirevic:2011bp} and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$
\cite{Faller:2013dwa}.

For this class of decays, we face the problem that $P$ has a given dependence on the dilepton helicity angle $\theta_1$
and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details on the angular distribution.)
However, on the level of decay amplitudes the dimeson system can have an arbitrary
large total angular momentum; only the third component of its angular momentum is restricted to $J^{\pi\pi}_z = -1,0,+1$.\\

We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(m)}$
with respect to the remaining angle $\theta_2$. Using the observables $P_k(\vec{\nu}=(q^2, k^2))$ as defined in \refapp{partial-waves}
the expansion reads
\begin{equation}
    P_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,m}} P_{k,j}(\vec{\nu}) p_{j}^{(m)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where $m=0$ for $n=1,2,6$ and $m=1$ for $n=3,4,5,7,8,9$. The angular observables defined as coefficients of this expansion
have the merit of a well-defined total angular momentum, and thus can be disentangled for
a derivation). As a consequence of the orthogonality of the Legendre polynomials, mismodelling or lack of modelling of higher
partial-wave observables \emph{does not affect} the method of moments as given in the previous section.
Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$P_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinear of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the S and P wave contributions. (For $B\to K\pi\ell^+\ell^-$ an initial study has been carried out for contributions up to the D wave \cite{Das:2014}).\\

@@

\subsection{Detector Acceptance Effects and Backgrounds}
\label{sec:systematics:acceptance}

\fred{If this is to be general, should explain what we mean by
  acceptance. Isn't it that experiments just model the detector well?
  The P. to observe an event in $\nuvec, \thvec$ might be very
  different from the P. to create an event in $\nuvec, \thvec$. All
  that's needed is a full forward model. I'm not sure that in
  \refeq{eq:PDFwithDE} we still have a normalized probability. Am I
  right to say that detector effects can be very much localized and
  nondifferentiable? Like some wires are inactive, or the region close
  to the beam has zero acceptance. In the latter case, modeling
  residual acceptance effects with smooth Legendre polynomials seems a
  bad idea.}

Thanks to accurate detector simulations and control studies, the experimental analyses have become
remarkably good at removing detector acceptance effects. In this section we therefore assume that
there are only a small, residual acceptance effects left that affect the events. In particular, we
assume that the latter can be modeled as a modified probability density,
\begin{equation}
    \label{eq:PDFwithDE}
    P(\nuvec,\thvec) \to P(\nuvec,\thvec) \times \Omega(\thvec),
\end{equation}
where $\Omega$ is independent of the non-angular kinematics $\nuvec$. The latter assumption is only
is invalid in general. However, if residual angular acceptance only weakly depends on $\vec{\nu}$, one
can always choose appropriate bin $\vec{\nu}$ so that $\Omega$ is approximately constant with respect
to $\vec{\nu}$. For such a case we parametrize
\begin{equation}
    \Omega_{m,l}(\thvec) = \prod_{i=1}^{m} \Big[1 + \sum_{j=1}^l \eps_i^{(j)} p_j(\cos\theta_i)\big]
\end{equation}
where $p_j(\cos\theta_i)$ denotes the $j$th Legendre polynomial. Since we only consider residual
acceptance effects, we expect $\eps_i^{(j)} \ll 1$. \footnote{\danny{what are reasonable values/upper limits on the $\eps$?}}\\

@@

For this type of distribution, we expand $\Omega_3$ to linear order in $\eps$,
\begin{equation}
    \Omega_{3,2}(\theta_1,\theta_2,\theta_3) = 1 + \sum_{i=1}^3 \sum_{j=1}^2 \eps_i^{(j)} p_j(\cos\theta_i) + \order{\eps^2}\,.
\end{equation}
This leads to a matrix-valued equation that relates the effiency-corrected estimators $\wwhat{P}_i$ to the uncorrected
estimators $\what{P}_j$,
\begin{equation}
    \label{eq:eff-corrected-est-btokstarll}
    \wwhat{P}_i = \sum_{j=1\text{s},\dots\,9} \big(\delta_{ij} + N_{ij}\big) \what{P}_j\,.
\end{equation}
We provide the matrix $N$ in \reftab{eff-corrected-est-btokstarll}.\\

>>>>

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{epstopdf}
%% Shortcuts %%
\newcommand{\dd}{\text{d}}
\newcommand{\refapp}[1]{app.~(\ref{app:#1})}
\newcommand{\refeq}[1]{eq.~(\ref{eq:#1})}
\newcommand{\refsec}[1]{sec.~(\ref{sec:#1})}
\newcommand{\reftab}[1]{tab.~(\ref{tab:#1})}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\nuvec}{\vec{\nu}}
\newcommand{\thvec}{\vec{\vartheta}}
\let\oldtheta\theta
\renewcommand{\theta}{\vartheta}
%\newcommand{\GeV}{\text{GeV}}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}
\let\eps\varepsilon
\newcommand{\aver}[1]{\langle #1 \rangle}
\newcommand{\est}[1]{\widehat{#1}}
\newcommand{\wwhat}[1]{\widehat{\widehat{#1}}}
\DeclareMathOperator{\cov}{Cov}
%\DeclareMathOperator{\skew}{Skew}
\DeclareMathOperator{\kurt}{Kurt}

@@

\captionsetup{compatibility=false}

@@

\preprint{SI-HEP-2014-17}
\title{Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays}
\author{Frederik Beaujean}
\email{frederik.beaujean@lmu.de}
\affiliation{C2PAP, Universe Cluster, Ludwig-Maximilians-Universit\"at M\"unchen, Garching, Germany}
\author{Marcin Chrzaszcz}
\email{mchrzasz@cern.ch}
\author{Nicola Serra}
\email{nicola.serra@cern.ch}
\affiliation{Physik-Institut, Universit\"at Z\"urich, Z\"urich, Switzerland}
\author{Danny van Dyk}
\email{vandyk@tp1.physik.uni-siegen.de}
\affiliation{Theoretische Physik 1, Naturwissenschaftlich-Technische Fakult\"at,
Universit\"at Siegen, Siegen, Germany}

\begin{abstract}
In this letter we study how to obtain a set of angular observables
$\{P_i\}$ that arise in a generic multi-body process without the need
to carry out a likelihood fit of an angular distribution to the
measured events. Instead, we present a method that only relies on
orthogonality of angular functions, and estimation of integrals
by means of Monte Carlo techniques. Beyond providing estimators
for central values and uncertainties, we provide means to determine
the correlation between all angular observables. We show that detector
acceptance effects can be accounted for in the analysis.
\end{abstract}

@@

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the rare decay FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
the method we describe in this letter is general, and it applies to arbitrary
decay or scattering processes. We find a previous work \cite{Dighe:1998vk}
advocates this method chiefly for the determination of angular observables in non-leptonic
$B$ decays, but also mentions the applicability to semileptonic decays. We improve
upon this previous work by studying correlations between the angular observables,
and sources of systematic uncertainties. Moreover, we will show that the method
at hand has two benefits over the usual approach based on likelihood fits.
\begin{enumerate}
    \item Likelihood fits do not always converge. \fred{Why? Mismodelling? Expand!}
    \item Using the method at hand, the joint probability distribution of the
        angular observables converges towards a multivariate normal distribution.
        This allows easy transfer of information from the experiment to interest
        theorists.
\end{enumerate}

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$f_i(\thvec)$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega f_i(\thvec) f_j(\thvec) \dd^N \theta = \delta_{ij}\,,
\end{equation}
with $\Omega$ representing the full angular phase space relevant to the process.
For the purpose of this letter, however, it suffices that the system
of angular functions $f_i(\thvec)$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays which we consider the most interesting
we list the transformations in a series of appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.\\

For particle decays, $P$ is generally expressed in terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\dd^{N+M}\Gamma}{\dd \nu_1 \dots \dd \nu_M\, \dd \theta_1 \dots \theta_N}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\dd^{N+M}\sigma}{\dd \nu_1 \dots \dd \nu_M\, \dd \theta_1 \dots \theta_N}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are invariant.\\

In the remainder of this letter we discuss how to obtain the angular
observables $S_i(\nuvec)$ in an experimental setup where each recorded
event is (approximately) distributed according to $P$.  We establish
the statistical basics in section \ref{sec:sample-based-det}. Section
\ref{sec:systematics} is dedicated to the impact of systematic effects
such as mismodeling the underlying physics or detector acceptance
effects. Numerical studies for one uni-angular and one triple-angular
distribution are provided in section \ref{sec:numerics}. In a series
of appendices we provide the orthonormal bases of angular functions
relevant to several rare $b$ decays.

@@

The orthonormality relations \refeq{def-ortho-rel} imply that a single angular observable $S_i$
can be projected out of the full PDF $P$ by means of
\begin{equation}
    \label{eq:det-Pi-analytical}
    S_i(\nuvec) = \int_{\Omega} \hat{f}_i(\thvec) P(\nuvec, \thvec) \dd^N \theta\,.
\end{equation}
where $\lbrace \hat{f}_i \rbrace$ denotes a \emph{dual basis} of angular function. In all generality
the bases $\lbrace f_i \rbrace$ and $\lbrace \hat{f}_i \rbrace$ are not identical. This is the
case for our selection of applications in appendices \ref{app:btokll} through \ref{app:lambdabtolambdall}.\\

It is sensible to refer to the angular observable $S_i$ as the \emph{$\hat{f}_i$-moment} of the PDF $P$.
We emphasize that a relation of type \refeq{det-Pi-analytical} holds for any combination of
scalar product and orthonormal basis of angular functions $\lbrace f_i \rbrace$; \ie, there is no unique
basis of angular functions. For the proof we refer to ref. \cite{Dighe:1998vk}.\\
In addition, integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int S_i(\nuvec) \dd^M \nu
    = \int \left[\int_{\Omega} \hat{f}_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\,.
\end{equation}
The remainder of this section describes the method of moments, in which we replace the
analytical integration \refeq{det-Pi-analytical} by Monte Carlo (MC) estimators.\\

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int g(x) P(x) \dd x
\end{equation}
can be approximated \cite{MCSM-2004} by an MC estimator $\est{E_P[g]}$
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1} g(x^{(k)})
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$; i.e.,
\begin{equation}
    x^{(k)} \sim P\,.
\end{equation}
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \hat{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
S_i\rangle$.
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} S_i(\nuvec) \dd^M \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\Omega} \hat{f}_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\\
    & = \int \left[\int_{\Omega} \hat{f}_i(\thvec) P(\nuvec,\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \dd^N\theta\right]\dd^M \nu\,,
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \hat{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This basically reduces the estimation of
$\aver{S_i}_{\vec{a},\vec{b}}$ to a (weighted) counting experiment. In
the limit $N \to \infty$, the random vector $( \dots,
\aver{S_i}_{\vec{a},\vec{b}}, \dots)$ follows a multivariate Gaussian
distribution \citeneeded\fred{Robert,Casella?}
with the mean vector estimated by \refeq{bin-importance}, and the covariance
estimated as
\begin{equation}
    \cov[\aver{S_i},\aver{S_j}]_{\vec{a},\vec{b}} \to \est{\cov}[\aver{S}_i, \aver{S}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\hat{f}_i(x^{(k)}) - \est{\langle S_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\hat{f}_j(x^{(k)}) - \widehat{\langle S_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly converge towards the true covariance matrix.
We propose to declare convergence once estimators for skewness and kurtosis are sufficiently small,
\begin{equation}
    \est{\operatorname{Skew}}[\aver{S_i},\aver{S_j}] \ll 1\,\,\text{ and }\,\,\est{\kurt}[\aver{S_i},\aver{S_j}] \ll 1\,.
\end{equation}

The convergence towards a normal distribution for the results allows to transfer correlation information to interested
theorists. In particular, this will allow to constrict predictive distributions of optimized observables.
(See e.g. \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf}
for definitions of such optimized observables in $B\to K^*\ell^+\ell^-$ decays, \cite{Faller:2013dwa} for application to
the decay $B\to \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$).

@@

In \refsec{sample-based-det}, we assume that the PDF $P$ describe the underlying physics accurately,
and that the experiment observes each event with perfect accuracy. In order to estimate systematic
uncertainties, we lift these assumptions.

@@

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
In this section, we take the interesting class of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$
as an example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B \to K\pi\ell^+\ell^-$, and
the $V_{ub}$ suppressed decay $B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
the small-width approximation and when assuming a pure $P$-wave, resonant final state. An extension to
$S-P$ interference has been studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012mb,Becirevic:2011bp},
and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study of $S$, $P$
and $D$ interference, see \cite{Das:2014sra}}.%
The PDFs for these decays are usually expressed in terms of one or a few partial waves
of the dimeson system. However, the angular momentum of the dimeson system is unbounded from above.
\\

For the selected class of decays, we face the problem that $P$ has a given dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, on the level of decay amplitudes the dimeson system can have an arbitrary
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$.\\

We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}, and we have
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion
have the merit of a well-defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section.\\

Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$S_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinear of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the S and P wave contributions.
(For $B\to K\pi\ell^+\ell^-$ a first study has been carried out where contributions up to the D wave are studied \cite{Das:2014sra}).\\

@@

\subsection{Detector Acceptance Effects}
\label{sec:systematics:acceptance}

\fred{Comment that in this section $\vec\nu$ and $\vec\theta$ are used after ``correction'' has been applied.}
\danny{Rethink order of this section.}
Thanks to accurate detector simulations and control studies, the experimental analyses have become
remarkably good at removing detector acceptance effects. In this section we therefore assume that
there are only a small, residual acceptance effects left that affect the events. In particular, we
assume that the latter can be modeled as a modified probability density,
\begin{equation}
    \label{eq:PDFwithDE}
    P(\nuvec,\thvec) \to P(\nuvec,\thvec) \times \Omega(\thvec),
\end{equation}
where $\Omega$ is independent of the non-angular kinematics $\nuvec$. The latter assumption is only
is invalid in general. However, if residual angular acceptance only weakly depends on $\vec{\nu}$, one
can always choose appropriate bin $\vec{\nu}$ so that $\Omega$ is approximately constant with respect
to $\vec{\nu}$. For such a case we parametrize
\begin{equation}
    \Omega_{m,l}(\thvec) = \prod_{i=1}^{m} \Big[1 + \sum_{j=1}^l \eps_i^{(j)} p_j(\cos\theta_i)\big]
\end{equation}
where $p_j(\cos\theta_i)$ denotes the $j$th Legendre polynomial. Since we only consider residual
acceptance effects, we expect $\eps_i^{(j)} \ll 1$. \footnote{\danny{what are reasonable values/upper limits on the $\eps$?}}\\

@@

For this type of distribution, we expand $\Omega_3$ to linear order in $\eps$,
\begin{equation}
    \Omega_{3,2}(\theta_1,\theta_2,\theta_3) = 1 + \sum_{i=1}^3 \sum_{j=1}^2 \eps_i^{(j)} p_j(\cos\theta_i) + \order{\eps^2}\,.
\end{equation}
This leads to a matrix-valued equation that relates the effiency-corrected estimators $\wwhat{S}_i$ to the uncorrected
estimators $\est{S}_j$,
\begin{equation}
    \label{eq:eff-corrected-est-btokstarll}
    \wwhat{S}_i = \sum_{j=1\text{s},\dots\,9} \big(\delta_{ij} + N_{ij}\big) \est{S}_j\,.
\end{equation}
We provide the matrix $N$ in \reftab{eff-corrected-est-btokstarll}.\\
